\chapter{Experimental Evidence}
\section{Efimov Trimers in Atomic Systems}
Ultracold atomic clouds provided the first staging ground for exploring Efimov physics and related few-body phenomena because of the ability to control atom-atom interactions by an external field. In these extremely dilute gases, with densities $n$, the probability for collisions can reach unity by tuning the s-wave scattering length to the unitary regime $n\abs{a^3}\gg1$. In experiments with trapped ultracold atomic and molecular gases of alkali atoms with tunable two-body interactions, the existance of Efimov trimers have been inferred from resonantly enhanced loss rates, either in atomic three-body recombination processes at negative $a$, when the Efimov state couples to the triatomic threshold, or in atom-dimer relaxation processes. 
\paragraph{Alkali atomic gases} 
The first observations of an Efimov state was in an ultracold gas of $\prescript{133}{}{\mathrm{Cs}}$ was reported in 2006 by the group of Grimm and coworkers \cite{Grimm:2006}. Later experiments have strengthen the evidence of these elusive states and in 2014 observations of the first exited Efimov state confirmed the theoretically predicted universal scaling properties of two succesive Efimov states \cite{Huang2014}. Efimov resonances in homogenic gases with other atomic species have been observed including $\prescript{85}{}{\mathrm{Rb}}$, $\prescript{39}{}{\mathrm{K}}$, $\prescript{7}{}{\mathrm{Li}}$, and three component fermi gases of $\prescript{6}{}{\mathrm{Li}}$. Also mixtures mixtures of $\prescript{41}{}{\mathrm{K}}$ and $\prescript{87}{}{\mathrm{Rb}}$ has been investigated.

\paragraph{Helium trimers}
Helium is a prime candidate to study Efimov physics with a more direct approach. The ground state of the helium trimer is not an Efimov state. However the first and second exited trimer have Efimov characteristics. Coulomb explosion imaging of helium trimers have revealed the geometric form of the trimers. \cite{Kunitski:2015qth}

\section{Efimov States in Nuclei }

\section{Four-body Recombination connected to Efimov Trimers} 
The Efimov scenario is even richer. In connection to an Efimov trimer, a pair of four-body states can form when a fourth atom approach. In accordance with the theoretical predictions, strong evidence for the existance of a pair of four-body states was provided in 2009 \cite{Grimm:2009}.